

```{r bikeshare Data}

getwd()
list.files()

wash <- read.csv('washington.csv')
names(wash)
ny <- read.csv('new_york_city.csv')
names(ny)
head(ny)
class(chi$Start.Time)
chi <- read.csv('chicago.csv')
names(chi)

```



install.packages('ggplot2')
library(ggplot2)

install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)




# Install lubridate package
install.packages("lubridate")
library(lubridate)

# ----------------------------------------------
# most common month, day, hour per city? 
# ----------------------------------------------

# hour -------------------

ny$start <- as.POSIXlt(ny$Start.Time, format="%Y-%m-%d%H:%M")
h_count_ny <- table(format(as.POSIXct(ny$start), format="%H"))
max_h_ny <- subset(h_count_ny, h_count_ny==max(h_count_ny))
max_h_ny

chi$start <- as.POSIXlt(chi$Start.Time, format="%Y-%m-%d%H:%M")
h_count_chi <- table(format(as.POSIXct(chi$start), format="%H"))
max_h_chi <- subset(h_count_chi, h_count_chi==max(h_count_chi))
max_h_chi

wash$start <- as.POSIXlt(wash$Start.Time, format="%Y-%m-%d%H:%M")
h_count_wash <- table(format(as.POSIXct(wash$start), format="%H"))
h_count_wash
max_h_wash <- subset(h_count_wash, h_count_wash==max(h_count_wash))
max_h_wash

# plot -------

max_hour <- rbind(table(format(as.POSIXct(ny$start), format="%H")), table(format(as.POSIXct(chi$start), format="%H")), table(format(as.POSIXct(wash$start), format="%H")))
barplot(max_hour, main="hours", ylab="count", col=terrain.colors(3), beside=T)
legend(2, 35000, legend=c("ny", "chi", "wash"), fill=terrain.colors(3))

  

# month -------------------

chi$Start.Time <- as.Date(chi$Start.Time)
month_count_chi <- table(months(chi$Start.Time))
max_month_chi <- subset(month_count_chi, month_count_chi==max(month_count_chi))
max_month_chi

wash$Start.Time <- as.Date(wash$Start.Time)
month_count_wash <- table(months(wash$Start.Time))
max_month_wash <- subset(month_count_wash, month_count_wash==max(month_count_wash))
max_month_wash

ny$Start.Time <- as.Date(ny$Start.Time)
month_count_ny <- table(months(ny$Start.Time))
max_month_ny <- subset(month_count_ny, month_count_ny==max(month_count_ny))
max_month_ny

# plot -------

max_month <- rbind(month_count_ny,month_count_chi, month_count_wash)
max_month
barplot(max_month, main="max_month", ylab="count", col=terrain.colors(3), beside=T)
legend(2, 90000, legend=c("ny", "chi", "wash"), fill=terrain.colors(3))



# day -------------------

ny$Start.Time <- as.Date(ny$Start.Time)
day_count_ny <- table(weekdays(ny$Start.Time))
max_day_ny <- subset(day_count_ny, day_count_ny==max(day_count_ny))
max_day_ny

wash$Start.Time <- as.Date(wash$Start.Time)
day_count_wash <- table(weekdays(wash$Start.Time))
max_day_wash <- subset(day_count_wash, day_count_wash==max(day_count_wash))
max_day_wash

chi$Start.Time <- as.Date(chi$Start.Time)
day_count_chi <- table(weekdays(chi$Start.Time))
max_day_chi <- subset(day_count_chi, day_count_chi==max(day_count_chi))
max_day_chi

# plot -------

max_day <- rbind(day_count_ny,day_count_chi, day_count_wash)
max_day
barplot(max_day, main="max_day", ylab="count", col=terrain.colors(3), beside=T)
legend(2, 60000, legend=c("ny", "chi", "wash"), fill=terrain.colors(3))


# ----------------------------------------------
# most common start, end station & trip per city? 
# ----------------------------------------------

# start station
subset(table(ny$Start.Station), table(ny$Start.Station)==max(table(ny$Start.Station)))
subset(table(wash$Start.Station), table(wash$Start.Station)==max(table(wash$Start.Station)))
subset(table(chi$Start.Station), table(chi$Start.Station)==max(table(chi$Start.Station)))

# end station
subset(table(ny$End.Station), table(ny$End.Station)==max(table(ny$End.Station)))
subset(table(wash$End.Station), table(wash$End.Station)==max(table(wash$End.Station)))
subset(table(chi$End.Station), table(chi$End.Station)==max(table(chi$End.Station)))

# trip
ny$trip <- paste(ny$Start.Station, ny$End.Station)
subset(table(ny$trip), table(ny$trip)==max(table(ny$trip)))

chi$trip <- paste(chi$Start.Station, chi$End.Station)
subset(table(chi$trip), table(chi$trip)==max(table(chi$trip)))

wash$trip <- paste(wash$Start.Station, wash$End.Station)
subset(table(wash$trip), table(wash$trip)==max(table(wash$trip)))


# ----------------------------------------------
# total/average trip duration per city? 
# ----------------------------------------------

total.trip <- sum(ny$Trip.Duration)
total.trip
avg.trip <- mean(ny$Trip.Duration)
avg.trip

total_y <- total.trip%/%(365*24*60*60)
total_y
rest_y <- total.trip%%(365*24*60*60)
rest_y

total_d <- rest_y%/%(60*60*24)
total_d
rest_d <- rest_y%%(60*60*24)
rest_d

total_h <- rest_d%/%(60*60)
total_h
rest_h <- rest_d%%(60*60)
rest_h

total_min <- rest_h%/%(60)
total_min
rest_sec <- rest_h%%(60)
rest_sec

#print("total trip: ", total_y,"years",total_d, "days", total_h, "hours", rest_sec, "sec")

total_y
total_d
total_h
total_min
rest_sec


# ----------------------------------------------
# user type, gender, dob-stats per city? 
# ----------------------------------------------

# counts of user type
ny_type <- data.frame(table(ny$User.Type))
chi_type <- data.frame(table(chi$User.Type))
wash_type <- data.frame(table(wash$User.Type))

barplot(table(ny$User.Type), table(chi$User.Type), table(wash$User.Type))

# plot ---------

ny_chi_merge <- merge.data.frame(ny_type, chi_type, by = "Var1" , all=T)
all_merge <- merge.data.frame(ny_chi_merge, wash_type, by = "Var1", all=T)
colnames(all_merge) <- c("user_type","ny","chi","wash")
all_merge


barplot(all_merge, main="user_type", ylab="count", col=terrain.colors(3), beside=T)
legend(2, 60000, legend=c("ny", "chi", "wash"), fill=terrain.colors(3))

barplot(table(ny$User.Type), )


ggplot(aes(x="Customer"), data=ny_type) + geom_bar()



# counts of gender
table(ny$Gender)
table(chi$Gender)

# most common bday
bd_count_ny <- table(ny$Birth.Year)
most_common_bd_ny <- subset(bd_count_ny, bd_count_ny==max(bd_count_ny))
most_common_bd_ny

bd_count_chi <- table(chi$Birth.Year)
most_common_bd_chi <- subset(bd_count_chi, bd_count_chi==max(bd_count_chi))
most_common_bd_chi

# most recent bday
max(na.omit(ny$Birth.Year))
max(na.omit(chi$Birth.Year))


# most earliest bday
min(na.omit(ny$Birth.Year))
min(na.omit(chi$Birth.Year))



```{r Faceting}

ggplot(aes(x=dob_day,y=..density..), data=pf)+
  geom_histogram(binwidth=1)+
  scale_x_continuous(breaks=1:31)+
  facet_wrap(~gender, ncol=3)

ggplot(aes(x=dob_day,y=..density..), data=pf)+
  geom_histogram(binwidth=1)+
  scale_x_continuous(breaks=1:31)+
  facet_grid(dob_month~gender)

```


### Friend Count
```{r Friend Count}

ggplot(aes(x=friend_count, y=..density..), data=pf)+
  geom_histogram(binwidth=1)+
  scale_x_continuous()
```
### Limiting the Axes + Exploring with Bin Width


```{r Limiting the Axes}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```


### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
ggplot(aes(x=friend_count),data=pf)+
  geom_histogram(binwidth = 25)+
  scale_x_continuous(limits=c(0,1000), breaks = seq(0,1000,50))+
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
ggplot(aes(x=friend_count),data=subset(pf, !is.na(gender)))+
  geom_histogram(binwidth = 25)+
  ggtitle('Histogram of Number of Friends Per Gender') +
  labs(x = "Number of Friends") + 
  scale_x_continuous(limits=c(0,1000), breaks = seq(0,1000,50))+
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)

by(pf$friend_count, pf$gender, summary)


```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

by(pf$www_likes, pf$gender, sum)



```


***

### Box Plots
Notes:

```{r Box Plots}


qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) +
        scale_x_continuous(limits=c(0,1000), breaks = seq(0,1000,50)) +
        facet_wrap(~gender)

# passing the geom argument allows us to change the plot type
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), geom = 'boxplot', ylim = c(0,100)) 


# passing the geom argument allows us to change the plot type
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), 
      geom = 'boxplot') + 
      coord_cartesian(ylim = c(0,100)) 

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Scatter Plots
Notes:

```{r scatter plots}

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point() +
  xlim(13,90)

summary(pf$age)


```


#### Overplottingr.

```{r Overplotting}

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_point(alpha=1/20) +
  xlim(13,90)



```
### Jitter Plots
Notes:

```{r Jitter plots}

ggplot(aes(x = age, y = friend_count), data = pf) +
  geom_jitter(alpha=1/20) +
  xlim(13,90)




```
### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(aes(x = age, y = friend_count), data = pf) +
  xlim(13,90)+
  geom_jitter(alpha=1/20,
              position = position_jitter(h=0),
              color='orange') +
  coord_trans(y='sqrt')+
  geom_line(stat='summary', fun.y=mean)+
  geom_line(stat='summary', fun.y=quantile,
            fun.args= list((probs = 0.1), (linetype=2), (color = 'blue')))




```

### Bivariate Plots
Notes:

```{r Bivariate Plots}

ggplot(aes(x = age, y = friend_count), data = pf) +
  xlim(13,90)+
  geom_jitter(alpha=1/20,
              position = position_jitter(h=0),
              color='orange') +
  coord_trans(y='sqrt')+
  geom_line(stat='summary', fun.y=mean)+
  geom_line(stat='summary', fun.y=quantile,
            fun.args= list((probs = 0.1), (linetype=2), (color = 'blue')))




```
